---
title: "AA-d13C fingerprinting - American Samoa Coral"
author: "Veronica Radice"
date: "10/16/2023"
output: html_document
---

```{r, global_options, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message=FALSE, tidy.opts=list(width.cutoff=70), tidy=TRUE)
```

```{r, echo=FALSE}
library(dplyr)
library(plyr)
library(factoextra)
library(FactoMineR)
library(ggplot2)
library(MASS)
```


#### Set working directory

```{r}
setwd("/Users/veronica/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/AmSam_isotopes/CSIA-AA_AmSam_coral")
```

### AA d13C data

```{r}
# changed d13C to d13C_AA because will merge with meta data sheet that has bulk isotope data
aa.d13C <- read.csv("AA_C_data_RadiceCoral_20230821_JDC.csv", header = TRUE)

aa.d13C[sapply(aa.d13C, is.character)] <- lapply(aa.d13C[sapply(aa.d13C, is.character)], as.factor)

head(aa.d13C)
```

### Summary - data set

Two different depths shallow and mesophotic *Montipora grisea*
```{r}
aa.d13C %>% 
  group_by(Group, Species) %>%
      dplyr::summarise(count = n())
```

### meta data

```{r}
meta <- read.csv("AmericanSamoa_Coral-isotopes_Master_2023-10.csv", header = TRUE)

meta[sapply(meta, is.character)] <- lapply(meta[sapply(meta, is.character)], as.factor)

head(meta)
```

```{r}
aa.d13C.meta <- meta %>% 
  inner_join(aa.d13C, by = c("Vial_Sample_ID", "Site", "Group", "Species"))

aa.d13C.meta <- droplevels(aa.d13C.meta)

dim(aa.d13C.meta)
```

##### select only the essential amino acids
```{r}
aa.d13C.ess <- aa.d13C.meta %>% dplyr::select(Group, Group2, Group3, Species, Depth, Vial_Sample_ID, Ile, Leu, Phe, Thr, Val)

#aa.d13C.ess <- aa.d13C.ess %>% 
#  filter(AA == "Ile" | AA == "Leu" | AA == "Phe" | AA == "Thr" | AA == "Val")

aa.d13C.ess <- droplevels(aa.d13C.ess)

head(aa.d13C.ess)
```

Select Coral Host (consumer) host data only
```{r}
host.aa.d13C.ess <- subset(aa.d13C.ess, Group == "Coral Host")
host.aa.d13C.ess <- droplevels(host.aa.d13C.ess)
```


########################################################################

## Sources

Relevant source (particulate organic sources from tropical coral reefs) CSIA-AA d13C values from the literature. Only included essential AA data.

```{r}
# had issues with importing csv
# can delete Lat and Long columns because read.csv() does not like 'degree' symbol
# also delete 'um' (micrometer) symbol to just um
# also delete columns with '/' (DOI, etc.) and '
# ** or otherwise use fileEncoding="latin1" and that worked
sources <- read.csv("Sources_EAA_literature.csv", header = TRUE, fileEncoding="latin1")

sources[sapply(sources, is.character)] <- lapply(sources[sapply(sources, is.character)], as.factor)

head(sources)
```

```{r}
levels(sources$Group)
```

*Some are phytoplankton samples from culture (Stahl et al. 2023)*

Filter out irrelevant data
```{r}
sources <- subset(sources, Group != "Macroalgae") # not a food source for coral
sources <- subset(sources, Notes != "15C") # temperature
sources <- subset(sources, Species != "Artemia salina") # also missing one AAess value, NA
sources <- subset(sources, Reference != "Shih et al. 2019 Microbial Ecology") # missing one AAess
#sources <- subset(sources, Reference != "Stahl et al. 2023 L&O") # culture phytoplankton
sources <- droplevels(sources)
```


```{r}
sources %>%
  group_by(Group, Environment) %>% 
    dplyr::summarise(count = n())
```

########################################################################

## Separate only source data

### Add our plankton and symbiont data to source data

```{r}
sources.Samoa <- subset(aa.d13C.meta, Group != "Coral Host")
sources.Samoa <- droplevels(sources.Samoa)
dim(sources.Samoa)
```

#### Merge our source data with literature sources

```{r}
sources.all <- full_join(sources, sources.Samoa, by = c("Group", "Species", "Group_Reference", "Group2", "Group3", "Group4", "Thr", "Ile", "Val", "Leu", "Phe", "Sample_ID", "Size_fraction"))
dim(sources.all)
```

```{r}
sources.all %>%
  group_by(Group) %>% 
    dplyr::summarise(count = n())
```

```{r}
sources.all %>%
  group_by(Group2) %>% 
    dplyr::summarise(count = n())
```


```{r}
sources.all %>%
  group_by(Group3) %>% 
    dplyr::summarise(count = n())
```

```{r}
sources.all$Group3 <- factor(sources.all$Group3, levels = c("Detritus", "Phytoplankton-Diatom", "Phytoplankton-Dinoflagellate", "Phytoplankton-field", "Symbiont", "Symbiont-Shallow-cave", "Zooplankton"))
```

```{r}
sources.all$Group4 <- factor(sources.all$Group4, levels = c("Detritus", "Phytoplankton-Cyanobacteria", "Phytoplankton-Diatom", "Phytoplankton-Dinoflagellate", "Phytoplankton-POM", "Phytoplankton-proxy", "Symbiont-Pocillopora-Palmyra", "Symbiont-Mesophotic", "Symbiont-Shallow", "Symbiont-Shallow-cave", "Zooplankton-Palmyra", "Zooplankton-AmSam"))
```


########################################################################

# Normalize source data *by the mean of all AAess from each source group/study*

```{r}
sources.all %>%
  group_by(Group_Reference) %>% 
    dplyr::summarise(count = n())
```

```{r}
# # need to use mean() instead of rowMeans()
# # because applying a function one row at a time, and so the x input is a 1-dimensional vector.
#
# sources.all <- sources.all %>% 
#   group_by(Group_Reference) %>% 
#     dplyr::mutate(EAAmean=rowMeans(c(Thr,Ile,Val,Phe,Leu)), na.rm = TRUE)
```

```{r}
sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(EAAmean=mean(c(Thr,Ile,Val,Phe,Leu)))
```

View column to see that EAAmean was averaged across Thr,Ile,Val,Phe,Leu for each Group_Reference
(confirmed with raw data)
```{r}
sources.all$EAAmean
```

Normalize for each AAess (EAA)
```{r}
sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(Thr_EAAn = Thr-EAAmean)

sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(Ile_EAAn = Ile-EAAmean)

sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(Val_EAAn = Val-EAAmean)

sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(Phe_EAAn = Phe-EAAmean)

sources.all <- sources.all %>% 
  group_by(Group_Reference) %>% 
    dplyr::mutate(Leu_EAAn = Leu-EAAmean)
```


```{r}
head(sources.all[,88:93])
```

########################################################################

# PCA - sources - exploratory

PCA analysis, this is just for a quick visual to see the stress vectors
```{r}
# by default, the function PCA() [in FactoMineR], standardizes the data automatically during the PCA; so you don’t need do this transformation before the PCA
pca.sources <- PCA(sources.all[,89:93], scale.unit = TRUE, graph = TRUE)
```

```{r}
pca.sources
```

# Eigenvalues / variances
```{r}
# examine the eigenvalues to determine the number of principal components to be considered

# get_eigenvalue(pca.sources): Extract the eigenvalues/variances of principal components

# eigenvalues measure the amount of variation retained by each principal component
# Eigenvalues are large for the first PCs and small for the subsequent PCs
# the first PCs corresponds to the directions with the maximum amount of variation in the data set

# An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the original variables in standardized data
# This is commonly used as a cutoff point for which PCs are retained
# This holds true only when the data are standardized.

eig.val <- get_eigenvalue(pca.sources)
eig.val
```


```{r}
pca.sources <- prcomp(sources.all[,89:93], scale. = TRUE, center = TRUE) #everything
summary(pca.sources)
```


#### Visualize the eigenvalues
##### SCREE plot - how many dimensions and what percent represent the data
```{r}
fviz_eig(pca.sources, addlabels = TRUE, ylim = c(0, 100))
```

```{r}
# get_pca_ind(res.pca), get_pca_var(res.pca): Extract the results for individuals and variables, respectively
# provides a list of matrices containing all the results for the active variables (coordinates, correlation between variables and axes, squared cosine and contributions)
var <- get_pca_var(pca.sources)
# Coordinates - var$coord: coordinates of variables to create a scatter plot
#head(var$coord)
# Cos2: quality on the factor map - var$cos2: quality of representation for variables on the factor map
# It’s calculated as the squared coordinates: var.cos2 = var.coord * var.coord.
#head(var$cos2)
```

```{r}
# Contributions to the principal components (in percentage) of the variables to the principal components
# The contribution of a variable (var) to a given principal component is (in percentage) : (var.cos2 * 100) / (total cos2 of the component)
head(var$contrib)
```

#### Quality of representation
```{r}
# The quality of representation of the variables on factor map is called cos2 (square cosine, squared coordinates)

# The closer a variable is to the circle of correlations, the better its representation on the factor map (and the more important it is to interpret these components)
# Variables that are closed to the center of the plot are less important for the first components

# library("corrplot")
# corrplot(var$cos2, is.corr=FALSE)

# plot variables
#fviz_pca_var(pca.sources, col.var = "black")

# Total cos2 of variables on Dim.1 and Dim.2
fviz_cos2(pca.sources, choice = "var", axes = 1:2)
```

#### Contributions of variables to PCs
```{r}
# contributions of variables in accounting for the variability in a given principal component are expressed in percentage.

# Variables that do not correlated with any PC or correlated with the last dimensions are variables with low contribution and might be removed to simplify the overall analysis.

# function corrplot() [corrplot] - highlight the most contributing variables for each dimension
# library("corrplot")
# corrplot(var$contrib, is.corr=FALSE)    

# Contributions of variables to PC1
fviz_contrib(pca.sources, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(pca.sources, choice = "var", axes = 2, top = 10)
# The total contribution to PC1 and PC2 is obtained with the following R code:
fviz_contrib(pca.sources, choice = "var", axes = 1:2, top = 10)

# red dashed line on the graph above indicates the expected average contribution
# If the contribution of the variables were uniform, the expected value would be 1/length(variables) = 1/10 = 10%
# For a given component, a variable with a contribution larger than this cutoff could be considered as important in contributing to the component
```

## PCA with ellipses
```{r}
# pdf('PCA_sources-all_Group2_95-CI.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group2, 
             #palette = c("black", "darkgreen"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Group2",
             mean.point = FALSE,
             #select.var = list(name = c("Thr_EAAn", "Ile_EAAn", "Val_EAAn", "Phe_EAAn", "Leu_EAAn")),
             #repel = TRUE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Environment, 
             #palette = c("black", "darkgreen"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Environment",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )
```

```{r}
fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group_Reference, 
             #palette = c("black", "darkgreen"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Group_Reference",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )
```

```{r}
# pdf('PCA_sources-all_Group3_50-CI_just4viz.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group3, 
             palette = c("dodgerblue2", "cadetblue4", "azure4", "black", "chartreuse3", "cyan2", "chocolate1"), #brown1
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.5,
             ellipse.type = "norm",
             legend.title = "Group3",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
# pdf('PCA_sources-all_Group3_95-CI.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group3, 
             palette = c("dodgerblue2", "cadetblue4", "azure4", "black", "chartreuse3", "cyan2", "chocolate1"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Group3",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
# pdf('PCA_sources-all_Group4_50-CI_just4viz.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group4, 
             palette = c("dodgerblue2", "grey", "azure4", "cornsilk3", "black", "cadetblue4", "chartreuse3", "darkolivegreen3", "aquamarine2", "cyan2", "chocolate1", "brown1"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.5,
             ellipse.type = "norm",
             legend.title = "Group4",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```


*Probably remove Symbiont-Pocillopora-Palmyra - quite distinct from AmSam symbiont data*

*Our AmSam symbiont data is distinct enough as is - clear separation by both reef depth/environment (shallow, mesophotic, and shallow cave environment) and genus (Leptoseris and Montipora)*

Separate our symbiont data by reef depth/environment (shallow, mesophotic, and shallow cave environment)

Can merge Zooplankton data from both in situ reef locations (AmSam and Palmyra)

Can merge Phytoplankton-POM (GFF 0.7 um), Phytoplankton-proxy (McMahon phyto data), and Phytoplankton-Cyanobacteria into one Phytoplankton category

Culture Phytoplankton-Dinoflagellate and culture Phytoplankton-Diatoms more distinct - keep separate?


```{r}
PC1<-pca.sources$x[,1]
PC2<-pca.sources$x[,2]
PCAloadings <- data.frame(Variables = rownames(pca.sources$rotation), pca.sources$rotation)
```


########################################################################

# PCA - Final sources

Without Symbiont-Pocillopora-Palmyra
```{r}
sources.all <- subset(sources.all, Group4 != "Symbiont-Pocillopora-Palmyra")
sources.all <- droplevels(sources.all)
dim(sources.all)
```

Relevel sources that are similar (same group)
```{r}
sources.all$Group4 <- revalue(sources.all$Group4, c("Zooplankton-AmSam" = "Zooplankton", "Zooplankton-Palmyra" = "Zooplankton", "Phytoplankton-POM" = "Phytoplankton", "Phytoplankton-proxy" = "Phytoplankton", "Phytoplankton-Cyanobacteria" = "Phytoplankton"))
```

```{r}
sources.all$Group4 <- factor(sources.all$Group4, levels = c("Detritus", "Phytoplankton", "Phytoplankton-Diatom", "Phytoplankton-Dinoflagellate", "Symbiont-Mesophotic", "Symbiont-Shallow", "Symbiont-Shallow-cave", "Zooplankton"))
```


# PCA - sources

PCA analysis, this is just for a quick visual to see the stress vectors
```{r}
# by default, the function PCA() [in FactoMineR], standardizes the data automatically during the PCA; so you don’t need do this transformation before the PCA
pca.sources <- PCA(sources.all[,89:93], scale.unit = TRUE, graph = TRUE)
```

```{r}
pca.sources
```

# Eigenvalues / variances
```{r}
# examine the eigenvalues to determine the number of principal components to be considered

# get_eigenvalue(pca.sources): Extract the eigenvalues/variances of principal components

# eigenvalues measure the amount of variation retained by each principal component
# Eigenvalues are large for the first PCs and small for the subsequent PCs
# the first PCs corresponds to the directions with the maximum amount of variation in the data set

# An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the original variables in standardized data
# This is commonly used as a cutoff point for which PCs are retained
# This holds true only when the data are standardized.

eig.val <- get_eigenvalue(pca.sources)
eig.val
```


```{r}
pca.sources <- prcomp(sources.all[,89:93], scale. = TRUE, center = TRUE) #everything
summary(pca.sources)
```


#### Visualize the eigenvalues
##### SCREE plot - how many dimensions and what percent represent the data
```{r}
fviz_eig(pca.sources, addlabels = TRUE, ylim = c(0, 100))
```

```{r}
# get_pca_ind(res.pca), get_pca_var(res.pca): Extract the results for individuals and variables, respectively
# provides a list of matrices containing all the results for the active variables (coordinates, correlation between variables and axes, squared cosine and contributions)
var <- get_pca_var(pca.sources)
# Coordinates - var$coord: coordinates of variables to create a scatter plot
#head(var$coord)
# Cos2: quality on the factor map - var$cos2: quality of representation for variables on the factor map
# It’s calculated as the squared coordinates: var.cos2 = var.coord * var.coord.
#head(var$cos2)
```

```{r}
# Contributions to the principal components (in percentage) of the variables to the principal components
# The contribution of a variable (var) to a given principal component is (in percentage) : (var.cos2 * 100) / (total cos2 of the component)
head(var$contrib)
```

#### Quality of representation
```{r}
# The quality of representation of the variables on factor map is called cos2 (square cosine, squared coordinates)

# The closer a variable is to the circle of correlations, the better its representation on the factor map (and the more important it is to interpret these components)
# Variables that are closed to the center of the plot are less important for the first components

# library("corrplot")
# corrplot(var$cos2, is.corr=FALSE)

# plot variables
#fviz_pca_var(pca.sources, col.var = "black")

# Total cos2 of variables on Dim.1 and Dim.2
fviz_cos2(pca.sources, choice = "var", axes = 1:2)
```

#### Contributions of variables to PCs
```{r}
# contributions of variables in accounting for the variability in a given principal component are expressed in percentage.

# Variables that do not correlated with any PC or correlated with the last dimensions are variables with low contribution and might be removed to simplify the overall analysis.

# function corrplot() [corrplot] - highlight the most contributing variables for each dimension
# library("corrplot")
# corrplot(var$contrib, is.corr=FALSE)    

# Contributions of variables to PC1
fviz_contrib(pca.sources, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(pca.sources, choice = "var", axes = 2, top = 10)
# The total contribution to PC1 and PC2 is obtained with the following R code:
fviz_contrib(pca.sources, choice = "var", axes = 1:2, top = 10)

# red dashed line on the graph above indicates the expected average contribution
# If the contribution of the variables were uniform, the expected value would be 1/length(variables) = 1/10 = 10%
# For a given component, a variable with a contribution larger than this cutoff could be considered as important in contributing to the component
```

## PCA with ellipses
```{r}
# pdf('PCA_sources-all_Group2_95-CI.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group2, 
             #palette = c("black", "darkgreen"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Group2",
             mean.point = FALSE,
             #select.var = list(name = c("Thr_EAAn", "Ile_EAAn", "Val_EAAn", "Phe_EAAn", "Leu_EAAn")),
             #repel = TRUE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Environment, 
             #palette = c("black", "darkgreen"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Environment",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )
```


```{r}
# pdf('PCA_sources-all_Group3_50-CI_just4viz.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group3, 
             palette = c("dodgerblue2", "cadetblue4", "azure4", "black", "chartreuse3", "cyan2", "chocolate1"), #brown1
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.5,
             ellipse.type = "norm",
             legend.title = "Group3",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
# pdf('PCA_sources-all_Group3_95-CI.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group3, 
             palette = c("dodgerblue2", "cadetblue4", "azure4", "black", "chartreuse3", "cyan2", "chocolate1"),
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.95,
             ellipse.type = "norm",
             legend.title = "Group3",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```

```{r}
# pdf('PCA_sources-all_Group4_50-CI_just4viz.pdf', width=5, height=4, paper='special')

fviz_pca_ind(pca.sources,
             axes = c(1,2),
             pointsize = 3,
             geom.ind = "point", # show points only (but not "text")
             col.ind = sources.all$Group4, 
             palette = c("dodgerblue2", "black", "azure4", "cadetblue4", "chartreuse3", "aquamarine2", "cyan2", "chocolate1", "brown1"), #"grey", "cornsilk3", "darkolivegreen3", 
             #palette = "Dark2",
             addEllipses = TRUE, # Concentration ellipses
             ellipse.level = 0.5,
             ellipse.type = "norm",
             legend.title = "Group4",
             mean.point = FALSE,
             ggtheme = theme_classic(base_size = 12),
             title = "" 
             )

#dev.off()
```


########################################################################

# Predicting isotopic fingerprints of primary producers

## Linear Discriminant Analysis (LDA)

Dimension reduction - find a linear combination of the predictors that gives maximum separation between the centers of the data while at the same time minimizing the variation within each group of data.

Uses MASS package.

### Normalize data by the mean

Coral host data
```{r}
#Normalize each essential AA
host.aa.d13C.ess$EAAmean = rowMeans(subset(host.aa.d13C.ess, select = c(Thr,Ile,Val,Leu,Phe)), na.rm = TRUE)
host.aa.d13C.ess$Thr_EAAn = host.aa.d13C.ess$Thr-host.aa.d13C.ess$EAAmean
host.aa.d13C.ess$Ile_EAAn = host.aa.d13C.ess$Ile-host.aa.d13C.ess$EAAmean
host.aa.d13C.ess$Val_EAAn = host.aa.d13C.ess$Val-host.aa.d13C.ess$EAAmean
host.aa.d13C.ess$Leu_EAAn = host.aa.d13C.ess$Leu-host.aa.d13C.ess$EAAmean
host.aa.d13C.ess$Phe_EAAn = host.aa.d13C.ess$Phe-host.aa.d13C.ess$EAAmean
```

# LDA with POM, Plankton, Symbionts, Detritus as sources - classify the animal fractions

#### create blank objects to fill with data
```{r}
symb.boot<-NULL
plkt.boot<-NULL
detritus.boot<-NULL
pom.boot<-NULL
#cyano.boot<-NULL
#diat.boot<-NULL
#dinofl.boot<-NULL
source.boot<-NULL

coral.host.class<-NULL

a<-NULL
b<-NULL
c<-NULL
d<-NULL

auto<-NULL
hetero<-NULL
a_mean<-NULL
h_mean<-NULL

dist<-NULL

coral.host.prop<-NULL
prop<-NULL
source.prop<-NULL
```


# LDA
```{r}
# The prior argument sets the prior probabilities of class membership. If unspecified, the class proportions for the training set are used. If present, the probabilities should be specified in the order of the factor levels.

## Calculate an error rate for our LDA model - using source dataset
# run an LDA with a jacknifing model fit, to look at error rate
# the line 'CV = TRUE' makes the LDA do jacknifed (leave-one-out cross-validation) model fit
group.lda.norm <- lda(Group ~ Ile_EAAn + Leu_EAAn + Phe_EAAn + Thr_EAAn + Val_EAAn, data = sources.all, CV = TRUE)
```

### classification of the LDA model to the sources
```{r}
# create a table which compares the classification of the LDA model to the sources
ct.prod.norm <- table(sources.all$Group, group.lda.norm$class)
ct.prod.norm
```

*Overall, Plankton and Symbiont did not classify particularly well*
- Plankton samples are highly heterogeneous, and likely contain phytoplankton as well (or otherwise plankton feeds on phytoplankton)
- Coral Symbionts are dinoflagellates and thus makes sense they group with POM dinoflagellates


Looking at more specific breakdown of Plankton (Phytoplankton vs. Zooplankton)
```{r}
ct.prod.norm.2<-table(sources.all$Group2, group.lda.norm$class)
ct.prod.norm.2
```

Further detail
```{r}
ct.prod.norm.3<-table(sources.all$Group3, group.lda.norm$class)
ct.prod.norm.3
```

Even more detail
```{r}
ct.prod.norm.4<-table(sources.all$Group4, group.lda.norm$class)
ct.prod.norm.4
```

```{r}
# total percent of samples correctly classified is the sum of the diagonal of this table
sum(diag(prop.table(ct.prod.norm)))
sum(diag(prop.table(ct.prod.norm.2)))
sum(diag(prop.table(ct.prod.norm.3)))
sum(diag(prop.table(ct.prod.norm.4)))
```


*In situ data with phytoplankton culture*
[1] 0.6435644
[1] 0.3168317
[1] 0.3267327

*In situ data only*
[1] 0.5405405
[1] 0.5675676
[1] 0.2567568

```{r}
# what % of each species is being correctly classified
diag(prop.table(ct.prod.norm, 1))
diag(prop.table(ct.prod.norm.2,1))
diag(prop.table(ct.prod.norm.3,1))
diag(prop.table(ct.prod.norm.4,1))
```


*In situ data with phytoplankton culture*
Detritus  Plankton       POM  Symbiont 
1.0000000 0.2666667 0.8076923 0.2500000 
1.00000000 0.07692308 0.60000000 0.25000000
1.0000000 0.0000000 1.0000000 0.3333333


*In situ data only*
Detritus   Plankton        POM   Symbiont 
1.00000000 0.06666667 0.56000000 0.55000000 
1.0000000 0.2400000 0.7333333 0.5500000
1.0000000 0.0000000 0.3636364 0.1250000


########################################################################



########################################################################


```{r}
sessionInfo()
```

